
\section*{Journal articles}
%

\textbf{Koelman, L.A.} and Lowery, M.M. (2019). Beta-band resonance and intrinsic oscillations in a biophysically detailed model of the subthalamic nucleus-globus pallidus network. Frontiers in computational neuroscience, 13, p.77.

\section*{Conference abstracts}
%
%
%

%
%

\textbf{Koelman, L.A.} and Lowery, M.M. (2017). Reduction of multi-compartmental subthalamic neuron model preserving somatodendritic interactions. Program No. 813.15 / WW32. 2017 Neuroscience Meeting Planner. Washington, DC: Society for Neuroscience, 2017. Online.\\

%

%

%

%
%
%

%
%

\noindent
\textbf{Koelman, L.A.} and Lowery, M.M. (2018). Model of phase locking in the subthalamo-pallidal loop in Parkinson's Disease. Program No. PII.12. 2018 Book of Abstracts. Dublin, Ireland: International Society of Electrophysiology and Kinesiology, 2018. Online.\\

%

%

%

%

%

%

%

%

%
%
%
%

\noindent
\textbf{Koelman, L.} and Lowery, M. (2018). Biophysical Model of Phase Locking in the Subthalamo-Pallidal and Pallido-Striatal Loops in Parkinson's Disease. Bernstein Conference 2018. doi: 10.12751/nncn.bc2018.0216 

%
%
%
%
%
%

%
%
%
%
%
%
%

%
%
%
%
%
%
%
%

%
%
%

%

%

%

%

%

\section*{Journal articles in preparation}

\noindent
\textbf{Koelman, L.} and Lowery, M. (2020). Network effects of high-frequency and phase-locked DBS in a biophysically detailed model of the STN-GPe loop. \textit{In Preparation}.\\

\noindent
\textbf{Koelman, L.} and Lowery, M. (2020). Requirements of reduced morphology neuron models for producing equivalent network behavior in a detailed model of the STN-GPe loop. \textit{In Preparation}.
%
%
